\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-12 tex/paper.tex
save

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Etiam commodo, enim
vitae facilisis pretium, ligula justo aliquet lectus, non interdum erat lorem et
nisi. Donec pharetra lectus in magna pellentesque vehicula. Praesent dapibus
lorem sed enim lacinia, a vestibulum sapien mattis. Sed ornare mollis aliquet.
Nulla tempor lacinia tortor, in rhoncus augue porta nec. Morbi sed convallis
nibh, eu hendrerit turpis. Curabitur sit amet rhoncus purus. Curabitur eget
magna lorem. Phasellus pretium nisi quis est tincidunt, faucibus vulputate augue
viverra. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin a urna a
ex egestas pulvinar. Curabitur facilisis a dui nec fringilla. Duis bibendum
fermentum quam. Donec eget urna eget mi placerat malesuada ut et lacus. Nulla ac
risus scelerisque, congue eros vitae, dignissim risus. Fusce purus ante,
vulputate sed magna et, varius imperdiet leo.

@@

@@

@@

    \caption{Architecture for the adversarial training of a binary classifier $f$ against a continuous parameter $Z$. \glnote{Shall we also illustrate the losses?}}
    \label{fig:architecture}
\end{figure*}

In this work, we repurpose adversarial training as a means to constraint the
predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
illustrated in Figure~\ref{fig:architecture}, we pit $f$ against an adversary
model $r := p_{\theta_r}(z | f(X;\theta_f))$ of parameters $\theta_r$ and
associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f))$
modeling the posterior probability density that $z$ parameterizes the sample $X$
observed through $f(\cdot;\theta_f)$. If $Z$ takes discrete values, then
$p_{\theta_r}$ can be represented as a probabilistic classifier $\mathbb{R}
\mapsto \mathbb{R}^{|{\cal Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$)
is the estimated probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly,
if $Z$ takes continuous values and if we assume some parameteric distribution
for $Z|f(X;\theta_f)$ (e.g., a truncated gaussian over a bounded support), then $p_{\theta_r}$ can be represented
as network whose output $j$ is the estimated value of the corresponding
parameter of that distribution (e.g., its mean and variance). As in
\citep{nix1994estimating}, the estimated probability density
$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$
under that distribution.  Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
then the corresponding correlation can be captured by $r$. By contrast, if
$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$ should
perform poorly and be close to random guessing. Training $f$ such that it
additionally minimizes the performance of $r$ therefore acts as a regularization
towards Eqn.~\ref{eqn:criterion}.

As for generative adversarial networks, we propose to
train $f$ and $r$ simultaneously, which we carry out by considering
the value function
\begin{equation}
    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_f, \theta_r)
\end{equation}
that we optimize by finding the saddle point $(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ such that
\begin{align}
    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \label{eqn:min_thetaf} \\
    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r) \label{eqn:max_thetar}.
\end{align}
Without loss of generality, the adversarial training procedure to obtain
$(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ is formally presented in
Algorithm~\ref{alg:adversarial-training} in the case of a binary classifier $f :
\mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
in Section~\ref{sec:theory}, ${\cal L}_f$ (resp. ${\cal L}_r$) is set to the
expected value of the
negative log-likelihood of $Y|X$ under $f$ (resp. of $Z|f(X;\theta_f)$ under
$r$). The optimization algorithm consists in using stochastic gradient descent
alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.

@@

\glnote{Discuss that we could equivalently require an intermediate representation of the data (as learned internally by $f$) to be pivotal, instead of requiring $f(X)$ to be pivotal.}

@@

>>>>

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Etiam commodo, enim
vitae facilisis pretium, ligula justo aliquet lectus, non interdum erat lorem et
nisi. Donec pharetra lectus in magna pellentesque vehicula. Praesent dapibus
lorem sed enim lacinia, a vestibulum sapien mattis. Sed ornare mollis aliquet.
Nulla tempor lacinia tortor, in rhoncus augue porta nec. Morbi sed convallis
nibh, eu hendrerit turpis. Curabitur sit amet rhoncus purus. Curabitur eget
magna lorem. Phasellus pretium nisi quis est tincidunt, faucibus vulputate augue
viverra. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin a urna a
ex egestas pulvinar.

@@

    \path[dashed,-] (5.25,-3) edge (5.25,-6.5);
    \node[below] at (5.25,-6.5) {${\cal L}_f(\theta_f)$};

@@

    \draw[dashed,-] (15.5,-5) -| (12.75,-6.5);
    \node[below] at (12.75,-6.5) {${\cal L}_r(\theta_f, \theta_r)$};

@@

    \caption{Architecture for the adversarial training of a binary classifier
    $f$ against a continuous parameter $Z$, assuming that $Z|f(X;\theta_f)$
    follows a truncated gaussian distribution.}
    \label{fig:architecture}
    \end{figure*}

In this work, we repurpose adversarial training as a means to constraint the
predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
illustrated in Figure~\ref{fig:architecture}, we pit $f$ against an adversary
model $r := p_{\theta_r}(z | f(X;\theta_f))$ of parameters $\theta_r$ and
associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f))$
modeling the posterior probability density that $z$ parameterizes the sample $X$
observed through $f(\cdot;\theta_f)$.
Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
then the corresponding correlation can be captured by $r$. By contrast, if
$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$ should
perform poorly and be close to random guessing. Training $f$ such that it
additionally minimizes the performance of $r$ therefore acts as a regularization
towards Eqn.~\ref{eqn:criterion}.

If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented e.g. as a
probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
$p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes continuous values and
if we assume some parameteric distribution for $Z|f(X;\theta_f)$ (e.g., a
truncated gaussian over a bounded support), then $p_{\theta_r}$ can be
represented e.g. as network whose output $j$ is the estimated value of the
corresponding parameter of that distribution (e.g., its mean and variance). As
in \citep{nix1994estimating}, the estimated probability density
$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$.
As further explained in the next section, let us note that the adversary $r$ may
take any form, i.e. it does need to be a neural network, as long as it exposes a
differentiable function $p_{\theta_r}(z|f(X;\theta_f))$ of sufficient capacity
to represent the true distribution within its bounded support.

As for generative adversarial networks, we propose to
train $f$ and $r$ simultaneously, which we carry out by considering
the value function
\begin{equation}
    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_f, \theta_r)
\end{equation}
that we optimize by finding the saddle point $(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ such that
\begin{align}
    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \label{eqn:min_thetaf} \\
    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r) \label{eqn:max_thetar}.
\end{align}
Without loss of generality, the adversarial training procedure to obtain
$(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ is formally presented in
Algorithm~\ref{alg:adversarial-training} in the case of a binary classifier $f :
\mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
in Section~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
expected value of the
negative log-likelihood of $Y|X$ under $f$ and of of $Z|f(X;\theta_f)$ under
$r$:
\begin{align}
    {\cal L}_f(\theta_f) &= \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
\end{align}
The optimization algorithm consists in using stochastic gradient descent
alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.

@@

@@

Interestingly, let us emphasize that these results hold using only the 1D output
of $f$ (in the case of binary classification) as input to the adversary. We
could similarly enforce an intermediate representation of the data to be
pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not necessary.
